---
title: "Hispanic Autocorrelation"
output: html_notebook
---
```{r}
library(spdep)
library(sf)     
library(dplyr)   
library(spData) 
library(ggplot2)
library(ggthemes)
setwd("C:\\Users\\laure\\OneDrive\\Desktop\\DIDA 425 Project\\nyc_shapefile")
```
```{r}
nyc <- st_read("NYC_Tract_ACS2008_12.shp")
```
```{r}
ggplot()+
  geom_sf(data = nyc, aes(fill=hispanic))+
  scale_fill_steps(
    name = "Hispanic Households",
    low = "lightsteelblue1",
    high = "tomato1",
    n.breaks = 4,
    show.limits = T)+
  theme_void()
```
```{r}
nyc_list <- nyc %>% 
  #first convert polygons to a neighbor object
  #R needs to know a unique ID for each neighborhood    
  #(cartodb_id)
  #Neighbors determined using Queen weights
  poly2nb(st_geometry(nyc)) %>% 
  #zero.policy just tells R not to remove 0 values
  nb2listw(zero.policy = TRUE)
```

```{r}
nyc_list %>% 
  moran.test(nyc$hispanic, ., zero.policy = TRUE)
```
The Moran statistic of .461 and a p value < 0.001 indicates a rejection of the null hypothesis.
This data was not generated using a spatially randomized process
```{r}
moran.plot(nyc$hispanic, 
           nyc_list, 
           zero.policy = TRUE, 
           labels = F,
           xlab = 'Number of Hispanic Households',
           ylab = 'Lagged Hispanic Households (of Neighbors)',
                   pch=20)
```
There is a positive relationship between the number of Hispanic Households and Hispanic Neighbors.
Areas with high numbers of Hispanic households tend to be near other areas with high numbers of Hispanic households

```{r}
lisaRslt <- localmoran(nyc$hispanic, nyc_list, 
            zero.policy = TRUE, na.action = na.omit)
head(lisaRslt)
```
```{r}
library(magrittr)
library(tidyr)
significanceLevel <- 0.05; # 95% confidence
meanVal <- mean(nyc$hispanic);

lisaRslt %<>% as_tibble() %>%
  set_colnames(c("Ii","E.Ii","Var.Ii","Z.Ii","Pr(z > 0)")) %>%
  #Create a new categorical value
  #This variable will call insignificant units "insignificant"
  #Outliers are determined by LI <0 - units higher than the mean are HL and units lower than the mean are LH
  #Clusters are units with LI >= 0, and classified as HH or LL based on whether they are above or below the mean
  mutate(coType = case_when(
  `Pr(z > 0)` > 0.05 ~ "Insignificant",
  `Pr(z > 0)` <= 0.05 & Ii >= 0 & nyc$hispanic >= meanVal ~ "HH",
  `Pr(z > 0)` <= 0.05 & Ii >= 0 & nyc$hispanic < meanVal ~ "LL",
  `Pr(z > 0)` <= 0.05 & Ii < 0 & nyc$hispanic >= meanVal ~ "HL",
  `Pr(z > 0)` <= 0.05 & Ii < 0 & nyc$hispanic < meanVal ~ "LH"))

# Now add this coType to original sf data
nyc$coType <- lisaRslt$coType %>% replace_na("Insignificant")

#Now we'll plot it!
ggplot(nyc) +
  geom_sf(aes(fill=coType),color = 'lightgrey') +
  scale_fill_manual(values = c('red','NA','blue','cyan'), name='Clusters & \nOutliers') +
  labs(title = "Number of Hispanic Families by Neighborhood")+
  theme_minimal()
```

