---
title: "Mapping City Bike Routes in NYC"
output: html_notebook
---
Used Dr.Haller's code to transform data into an sf object. Then combined the citibike and census tract data.
Plans:
- Map where the stations are located on top of the percentage of black/hispanic nyc residents
- Map where the stations are located on top of the percentage of poor areas of nyc
  - median income
  - median house value
  - unemployment rate
  - population over 18
  - median age male vs median age female
- Map where color coordinate who is a member and who is not in these areas as well
- Which boroughs are the most accessible? 

```{r}
library(sf)
library(ggplot2)
library(dplyr)
library(tidyverse)

#load in the bike routes first
setwd("C:/Users/13157/Downloads/CitiBike_project/nyc_bikes")
nyc_shape <- st_read("geo_export_bbe21116-30ae-47df-a002-919a94088bf7.shp")

#This is already a spatial object (called SF objects in R), so we can go ahead and map it right away
#Some data will need to be transformed to a spatial object with a coordinate reference system to be mapped
#ggplot()+
  #geom_sf(nyc_shape, mapping=aes())

#Check the coordinate reference system (CRS) - this is using the WGS84 projection, which is pretty standard
nyc_shape

```
```{r}
setwd("C:/Users/13157/Downloads/CitiBike_project/Borough_Boundaries")
nyc_boro <- st_read("geo_export_4e5aa31d-3ac8-4a6f-9077-d75efb6b0232.shp")

# Plot the bike routes and boroughs
ggplot() +
  geom_sf(data = nyc_shape, color = "red") + # Plot bike routes
  geom_sf(data = nyc_boro, fill = "transparent", color = "black", lwd = 1) + # Plot boroughs
  theme_minimal()

nyc_boro 
```

```{r}
#now load in NYC basemap
setwd("C:/Users/13157/Downloads/CitiBike_project/nyctract_acs")
nyc <- st_read("NYC_Tract_ACS2008_12.shp")
#The CRS of both maps needs to match! We use EPSG codes to set the CRS. This one is the same CRS as the other map layer
nyc <- st_set_crs(nyc, 'EPSG:4326')


#Now we can map it
#ggplot() +
  #geom_sf(nyc, mapping=aes(), fill = "white") +
  #geom_sf(data = nyc_shape, color = "red") + # Plot bike routes
  #geom_sf(data = nyc_boro, fill = "transparent", color = "black", lwd = 1) + # Plot boroughs
  #theme_minimal()

nyc
```

```{r}
#We can also plot the City Bikes trip data! Let's load it in
setwd("C:/Users/13157/Downloads/CitiBike_project")
trip_data <- read.csv("202306-citibike-tripdata.csv")

#get the data ready to map
#trip_data <- trip_data %>%  
  #filter(start_lng < 0) %>% 
  #I need to transform my lat/long coordinates into an sf object 
  #st_as_sf(coords = c("start_lng", "start_lat"), crs = 4326)
```

```{r}
# we need to group the stations by where the points are in each census tract

# using tutorial from https://rstudio-pubs-static.s3.amazonaws.com/1113537_5c12778f1092412cbe4913b82618da98.html

citi_lat_lng_6_23 <- trip_data %>% select(c("start_lat","start_lng")) 
```

```{r}

write.csv(citi_lat_lng_6_23, file = "C:/Users/13157/Downloads/CitiBike_project/citi_lat_lng_6_23.csv", row.names = FALSE)
 # save as csv so we don't have to keep loading in the entire dataset, save file path 
```

```{r}
# load in citi lat long from 6/23 now

setwd("C:/Users/13157/Downloads/CitiBike_project")

citi_lat_lng_6_23 <- read.csv("citi_lat_lng_6_23.csv")
```


```{r}

citi_points <- st_as_sf(citi_lat_lng_6_23, coords = c("start_lng", "start_lat")) # convert citi bike lat/long into sf object points

st_crs(citi_points) <- st_crs(nyc) # set crs same as nyc census tract

```

```{r}
citi_points
```


```{r}
#takes a long time to load, be patient

citi_nyc_joined <- st_join(citi_points, nyc, join = st_within) # join citi bike points with census tract
```


```{r}
# count how many points are in each tract

census_tract_points_count <- count(as_tibble(citi_nyc_joined), tractid) %>%
                                   arrange(desc(n)) %>% print()

```

```{r}
# spatial join nyc and census tract points count 
spatial_count_result <- left_join(nyc, census_tract_points_count, by = c("tractid" = "tractid"))
```

```{r}
# save the result as a shapefile, change to your own directory

st_write(spatial_count_result, "C:/Users/13157/Downloads/CitiBike_project/spatial_citibike_count_result.shp")
```

```{r}
# remember "n" is the number of people who used a station within that census tract
library(mapview)
mapview(spatial_count_result, zcol = "n", layer.name = "african")
```








